### 2021-07-21 paper/paper.tex
WIP: main design choices

@@

\section*{TODO}

\begin{itemize}
  \item Basic design of \trixi and its capabilities
  \item PID comparison of Fluxo and \trixi
  \item Discussion: Julia for numerical simulation science
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks?
  \item Shall we re-create the schematic overview in tikz?
\end{itemize}

@@

TODO %TODO

@@

\subsection{Review of the main design criteria}

TODO %TODO

\section{Performance comparison with FLUXO}

@@

TODO: Andrew, please describe the setups that you used. %TODO

@@

TODO: Plot of this comparison with the \lstinline{TreeMesh}?

@@

TODO %TODO
\begin{itemize}
  \item Packages can really be used together at low (or no) cost
  \item Reproducibility
  \item Good software development practices (CI, tests, version management)
  \item Relatively easy to integrate existing C libraries (p4est, HDF5)
  \item more nice features
  \item Can be extended: Multi-physics with 350 lines of code
  \item Compilation time can be annoying but also allows cool features and performance \cite{schlottkelakemper2021purely}
  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
\end{itemize}

@@

Funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation)
under Germany's Excellence Strategy EXC 2044-390685587, Mathematics Münster:
Dynamics-Geometry-Structure.

TODO: Add your funding sources (if any) %TODO

@@

>>>>

\newcommand{\todo}[1]{{\color{red}#1}}

@@

@@

\todo{TODO} %TODO

@@

\subsection{Review of the main design choices}

The design of \trixi was guided by extensibility, ease of use, and efficiency
(in this particular order). Thus, it will likely be possible to make (some parts)
even faster while sacrificing simplicity. Nevertheless, \trixi is already quite
fast, even compared to established high-performance open source software,
as described in details in Section~\ref{sec:performance-comparison}. In this
section, we focus mostly on the first two goals.

\subsubsection{\trixi is a library}

The most important design decision was to create \trixi as a library. In the
context of simulation-focused scientific computing, there are a lot of (open
or closed source) codes that are designed as monolithic applications that can
simulate a specific setup. They can often be configured in limited amounts by
compile time options and parameter files. In contrast, users of \trixi do not
need to compile anything or learn the syntax of parameter files. They just load
the library and write their simulation setups in Julia. Such scripts describing
a simulation setup for \trixi are called elixirs. As of version v0.3.43, \trixi
comes with \ca 200 example elixirs. A basic elixir could look as follows.

\begin{lstlisting}[language = Julia]
# Load all required libraries
using Trixi, OrdinaryDiffEq, Plots

# Set up a 2D linear advection problem with
# advection velocity `a`
a = (1.0, 1.0)
equations = LinearScalarAdvectionEquation2D(a)

# Choose an initial condition coming with Trixi.jl
ic = initial_condition_convergence_test

# Set up a discontinuous Galerkin spectral element
# method with given polynomial degree and surface
# flux implemented in Trixi.jl
solver = DGSEM(polydeg=3,
               surface_flux=flux_lax_friedrichs)

# Create a uniformely refined mesh with periodic
# boundary conditions in a square domain
coordinates_min = (-1.0, -1.0) # lower left
coordinates_max = ( 1.0,  1.0) # upper right
mesh = TreeMesh(coordinates_min, coordinates_max,
                initial_refinement_level=4,
                n_cells_max=10^5, periodicity=true)

# Create semidiscretization with all spatial
# discretization-related components
semi = SemidiscretizationHyperbolic(
  mesh, equations, ic, solver)

# Create an ODE problem from the semidiscretization
# with time span from 0.0 to 1.0
ode = semidiscretize(semi, (0.0, 1.0))

# Evolve the ODE problem in time using `solve` from
# OrdinaryDiffEq with adaptive time stepping
sol = solve(ode, RDPK3SpFSAL49(), abstol=1.0e-6,
  reltol=1.0e-6, save_everystep=false)

# Plot the numerical solution at the final time
plot(sol)
\end{lstlisting}

\subsubsection{Functions are pure Julia functions}

Since functions are first-class citizens in Julia, they can be passed around
and used efficiently. Thus, everything that acts like a function can be used
in \trixi, \eg, to define initial/boundary conditions. While such a flexibility
is not too uncommon, two-point numerical fluxes are also just functions with a
specified signature. Thus, users can write their own numerical fluxes in their
own code and they will work and be as efficient as if they were implemented
directly in \trixi.

\subsubsection{Ingredients use common interfaces and can be exchanged}

Abstractions such as the \lstinline{mesh} and the \lstinline{equations} are
based on a common interface. Using multiple dispatch in Julia,

\todo{TODO} %TODO

\subsubsection{New physics can be specified with minimal effort}

Instead of providing only the very basic discretization ingredients such as
other open source frameworks for PDEs, \trixi includes also some widely used
physical parts to make it easier to use.

\todo{TODO} %TODO

\subsubsection{Adding new functionality does not require changing existing code}

\todo{TODO} %TODO

\subsubsection{There is no spooky action at a distance}

\todo{TODO} %TODO

\section{Performance comparison with FLUXO}
\label{sec:performance-comparison}

@@

\todo{TODO: Andrew, please describe the setups that you used} %TODO

@@

\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}

@@

\todo{TODO} %TODO
\begin{itemize}
  \item Packages can really be used together at low (or no) cost
  \item Reproducibility
  \item Good software development practices (CI, tests, version management)
  \item Relatively easy to integrate existing C libraries (p4est, HDF5)
  \item more nice features
  \item Can be extended: Multi-physics with 350 lines of code
  \item Compilation time can be annoying but also allows cool features and performance \cite{schlottkelakemper2021purely}
  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
\end{itemize}

@@

We would like to thank all other contributors to and users of \trixi for their
help, encouragement, and discussions.
Funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation)
under Germany's Excellence Strategy EXC 2044-390685587, Mathematics Münster:
Dynamics-Geometry-Structure.

\todo{TODO: Add your funding sources (if any)} %TODO

@@

\todo{ % TODO
\begin{itemize}
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{schlottkelakemper2021purely,revels2016forward,ranocha2021sbp}
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks?
  \item Shall we re-create the schematic overview in tikz?
  \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
\end{itemize}}
